---
title: "Figures"
author: "Dr. Panagiotis Chouvardas"
date: "2024-12-02"
output: 
  github_document:
    toc: TRUE
---

## Pre-processing

### Loading libraries

```{r lib_load, warning=FALSE, message=FALSE}
library(ggplot2)
library(pals)
library(factoextra)
library(reshape)
library(ggpubr)
library(matrixStats)
library(tidyverse)
library(decoupleR)
library(rcartocolor)
`%!in%` = Negate(`%in%`)
library(BiocGenerics)
library(DESeq2)
library(ggrepel)
library(rcartocolor)
library(colourvalues)
library(readxl)
library(ggh4x)
library(tidymodels)


```

### Reading master file and creating gleason groups

```{r reading_master, warning=FALSE, message=FALSE}
cores <- read_xlsx("masterfile_GS_TC_allcores.xlsx")
cores$Sample <- paste0(cores$`Lab Code`, cores$Core)
cores$condition <- "Tumor"
cores$condition[which(cores$`Pathology evaluation` == "normal")] <- "Benign"
cores$GS_GRP <- "Benign"
cores$GS_GRP[which(cores$`Pathology evaluation` %in% c("3+3"))] <- "Group 1"
cores$GS_GRP[which(cores$`Pathology evaluation` %in% c("3+4"))] <- "Group 2"
cores$GS_GRP[which(cores$`Pathology evaluation` %in% c("4+3"))] <- "Group 3"
cores$GS_GRP[which(cores$`Pathology evaluation` %in% c("4+4"))] <- "Group 4"
cores$GS_GRP[which(cores$`Pathology evaluation` %in% c("4+5","5+4", "5+5"))] <- "Group 5"
```

## Figure 1

### 1C. Cohort.

```{r 1C, warning=FALSE, message=FALSE}
cores_gs2 <- melt(table(cores$`Lab Code`, cores$GS_GRP))
colnames(cores_gs2) <- c("Sample", "GS_GRP", "count")
cores_gs2$Sample <- factor(cores_gs2$Sample)
ggplot(cores_gs2, aes(x=Sample, y=count, fill=GS_GRP)) + geom_bar(stat="identity", position="fill", col="black") +
  theme_classic2() + 
  scale_fill_manual(values = color_values(16:1,"piyg")[c(2,9,10,12,14,16)]) + 
  theme(legend.position = "top") + ylab("Fraction (n = 4/sample)") + theme(axis.text = element_text(size = 7))  
```

### 1D. Organoid formation

```{r 1D, warning=FALSE, message=FALSE, fig.width = 3}
rate <- data.frame(Success = c("Yes", "Yes", "No", "No"), Core = c("Tumor","Benign", "Tumor", "Benign"),
                   count = c(41, 29, 15, 11))

strip <- strip_themed(background_x = elem_list_rect(fill = color_values(1:16,"piyg")[c(16,2)]), text_x = element_text(color="white"))

ggplot(rate, aes(x=Core, y=count, fill=Success)) + geom_bar(stat = "identity", position = "fill", col="black")+ 
  facet_wrap2(~Core, nrow = 2, scales = "free", strip = strip) + theme_classic2() + scale_fill_manual(values = rev(c("gray", "white"))) +
  geom_text(aes(label=count),position = position_fill(vjust = 0.5)) + theme(legend.position = "top") +
  ylab("Fraction") + theme(axis.text.x = element_blank()) + theme(axis.title.x = element_blank()) +
  guides(fill = guide_legend(title.position = "top", title.hjust = 0.5))
```

### 1E. Organoid Morphology

```{r 1E, warning=FALSE, message=FALSE}
df2 <- cores
df2[,9:11] <- df2[,9:11]/rowSums(df2[,9:11])
df2 <- as.data.frame(df2)
df2 <- melt(as.data.frame(df2), measure.vars = c("Solid", "Hollow", "Mixed"))
ggplot(df2, aes(x=condition, y=value, fill=condition)) + geom_boxplot() + geom_jitter(size=0.5) +
  facet_wrap(~variable) + stat_compare_means(aes(label = paste0("p = ", after_stat(p.format))), hjust=-0.5, size=2) +
  scale_fill_manual(values = color_values(1:11,"piyg")[c(10,2)]) + theme_classic2() +
  ylab("Fraction") + theme(legend.position = "none") +
  theme(axis.text = element_text(size=6)) + theme(axis.title = element_text(size=10)) +theme(axis.title.x = element_blank()) 
```

### 1F. Organoid counts

```{r 1F, warning=FALSE, message=FALSE}
df2 <- cores
df2 <- as.data.frame(df2)
df2 <- melt(as.data.frame(df2), measure.vars = c("Solid", "Hollow", "Mixed"))
ggplot(df2, aes(x=condition, y=value, fill=condition)) + geom_boxplot() + geom_jitter(size=0.5) +
  stat_compare_means(aes(label = paste0("p = ", after_stat(p.format))), hjust=-0.5, size=2) +
  scale_fill_manual(values = color_values(1:11,"piyg")[c(10,2)]) + theme_classic2() +
  ylab("Total Organoid Count") + theme(legend.position = "none") +
  theme(axis.text = element_text(size=6)) + theme(axis.title = element_text(size=10)) +theme(axis.title.x = element_blank()) 
```

## Sup. Figure 1

### S1A. Patient cohort GS

```{r S1A, warning=FALSE, message=FALSE}
clin <- read_xlsx("Clinical.xlsx", sheet = "Sheet1")
clin$GS <- paste0("(", clin$Path_GS1, "+", clin$Path_GS2, ")")
test_samples <- gsub("PCa", "", clin$SAMPLE_NR)
test_samples <- gsub("_P", "", test_samples)
test_samples <- gsub("_", "", test_samples)
clin$sample_new <- test_samples
clin <- filter(clin, sample_new %in% cores$`Lab Code`)
clin$GS_GRP <- "Group 1"
clin$GS <- gsub("\\(", "", clin$GS)
clin$GS <- gsub("\\)", "", clin$GS)
clin$GS_GRP[which(clin$GS %in% c("3+3"))] <- "Group 1"
clin$GS_GRP[which(clin$GS %in% c("3+4"))] <- "Group 2"
clin$GS_GRP[which(clin$GS %in% c("4+3"))] <- "Group 3"
clin$GS_GRP[which(clin$GS %in% c("4+4"))] <- "Group 4"
clin$GS_GRP[which(clin$GS %in% c("4+5","5+4", "5+5"))] <- "Group 5"
clin2 <- melt(table(clin$GS_GRP))
clin2$Var.1 <- factor(clin2$Var.1)

ggplot(clin2[which(clin2$Var.1 != 0),], aes(x=Var.1, y=value, fill=Var.1)) + geom_bar(stat="identity", col="black") +
  theme_classic2() + xlab("GS Group") + scale_fill_manual(values = color_values(16:1,"piyg")[c(9,10,12,14,16)]) +
  geom_text(aes(label=value),position = position_fill(vjust = 0.5), size=2) + ylab("# of Patients") + theme(legend.position = "none")
```

### S1B. Tumor content per core

```{r S1B, warning=FALSE, message=FALSE}
cores$`%tumor_content` <- gsub("<","",cores$`%tumor_content`)
cores2 <- filter(cores, `Pathology evaluation` != "normal")
cores2$`%tumor_content` <- as.numeric(cores2$`%tumor_content`)

ggplot(cores2, aes(x=GS_GRP, y=`%tumor_content`, fill=GS_GRP)) + 
  geom_boxplot() + geom_jitter() +ylab("Tumor Content (%)") +
  theme_classic2() + theme(legend.position = "none") +
  xlab("GS") + scale_fill_manual(values = color_values(16:1,"piyg")[c(9,10,12,14,16)]) 
```

## RNA seq data pre-processing

```{r RNA, warning=FALSE, message=FALSE}
my_pal <- c(color_values(1:11,"piyg")[c(10)], "lightgray", color_values(1:11,"piyg")[c(2)])
cores <- read_xlsx("masterfile_GS_TC_allcores.xlsx")
cores$Sample <- paste0(cores$`Lab Code`, cores$Core)
cores$condition <- "Tumor"
cores$condition[which(cores$`Pathology evaluation` == "normal")] <- "Benign"
df <- read.table("Processed.counts.csv", header = T, sep=",", row.names = 1)
df <- df[which(!duplicated(df$gene_name)),]
rownames(df) <- df$gene_name
df <- df[,-1]
colnames(df) <- gsub("PCa_","",colnames(df))
colnames(df) <- gsub("_TS","",colnames(df))
colnames(df) <- gsub("_","",colnames(df))

design <- data.frame(Sample = colnames(df))
design <- cbind.data.frame(design, Patient = substr(design$Sample, 1, 3))
rownames(design) <- design$Sample

design2 <- merge(design, cores, by="Sample")
rownames(design2) <- design2$Sample
design2 <- design2[colnames(df),]

dds <- DESeqDataSetFromMatrix(countData=df, 
                              colData=design2, 
                              design=~condition, tidy = FALSE)

design2 <- design
samples_to_keep <- names(table(design2$Patient))[which(table(design2$Patient) > 1)]
design2 <- filter(design2, Patient %in% samples_to_keep)

df2 <- df[,colnames(df) %in% rownames(design2)]
design2 <- design2[colnames(df2),]


dds <- DESeqDataSetFromMatrix(countData=df2, 
                              colData=design2, 
                              design=~Patient, tidy = FALSE)
dds <- dds[(rowSums(counts(dds) > 3) > 3 ),]
dds <- DESeq(dds)

vsd <- vst(dds)
df_norm <- t(assay(vsd))

rna.pr <- prcomp(df_norm, center = T, scale = TRUE)
pca_df <- data.frame(rna.pr$x[,1:2], Sample = rownames(rna.pr$x[,1:2]))
pca_df <- merge(pca_df, design2, by="Sample")
pca_df <- merge(pca_df, cores, by="Sample")

```

## Sup. Figure 6

### S6A. PCA tumor vs benign

```{r S6A, warning=FALSE, message=FALSE}
ggplot(pca_df, aes(x = PC1, y = PC2, col=condition, fill=condition)) + geom_point(size=4, shape=21, col="black") +
  theme_classic2() + theme(legend.position = "top", legend.title = element_blank()) +
  scale_color_manual(values = my_pal[c(1,3)]) +
  scale_fill_manual(values = my_pal[c(1,3)]) +
  ggforce::geom_mark_ellipse(aes(group=condition))
```

### S6B. DE/Volcano plot tumor vs benign

```{r S6B, warning=FALSE, message=FALSE}
design <- data.frame(Sample = colnames(df))
design <- cbind.data.frame(design, Patient = substr(design$Sample, 1, 3))
rownames(design) <- design$Sample

design2 <- merge(design, cores, by="Sample")
rownames(design2) <- design2$Sample
samples_to_keep <- names(table(design2$Patient))[which(table(design2$Patient) > 1)]
design2 <- filter(design2, Patient %in% samples_to_keep)
df2 <- df[,colnames(df) %in% rownames(design2)]
design2 <- design2[colnames(df2),]

dds <- DESeqDataSetFromMatrix(countData=df2, 
                              colData=design2, 
                              design=~condition, tidy = FALSE)
dds <- dds[(rowSums(counts(dds) > 3) > 3 ),]
dds <- DESeq(dds)

res <- results(dds, contrast = c("condition", "Tumor", "Benign"))
res <- as.data.frame(res)
res <- res[order(res$padj, decreasing = F),]
res <- cbind.data.frame(res, Gene_name = rownames(res))
res$log2FoldChange[which(res$log2FoldChange < -7)] <- -7
res$log2FoldChange[which(res$log2FoldChange > 7)] <- 7
res$DE <- "ns"
res$DE[which((res$log2FoldChange < -1)&(res$padj<0.05))] <- "Downregulated"
res$DE[which((res$log2FoldChange > 1)&(res$padj<0.05))] <- "Upregulated"
res <- na.omit(res)
ggplot(res, aes(x=log2FoldChange, y=-log10(padj), col=DE)) + geom_point() + theme_classic2() + 
  geom_vline(xintercept = 1, linetype= "dashed", col="gray") +
  geom_vline(xintercept = -1, linetype= "dashed", col="gray") +
  geom_hline(yintercept = -log10(0.05), linetype= "dashed", col="gray") +
  xlab("log2FoldChange\n(Tumor vs Benign)") + scale_color_manual(values = my_pal) +
  theme(legend.position = "top") + theme(legend.title = element_blank()) +
  annotate("text", x=-5, y=30, label = paste0("(",table(res$DE)[1], ")"), col=my_pal[1])+
  annotate("text", x=5, y=30, label = paste0("(",table(res$DE)[3], ")"), col=my_pal[3]) +
  xlim(-max(abs(res$log2FoldChange)), max(abs(res$log2FoldChange)))
```

### S6C. Master regulator analysis, tumor vs benign

```{r S6C, warning=FALSE, message=FALSE}
net <- get_collectri()
degs <- filter(res, padj < 0.05 & abs(log2FoldChange) > 1)
rownames(degs) <- degs$Gene_name
contrast_acts <- run_viper(degs[,"stat",drop=F], net, minsize = 10)

ggplot(contrast_acts[order(abs(contrast_acts$score), decreasing = T)[1:10],], aes(x = reorder(source, score), y = score)) + 
  geom_bar(aes(fill = score), stat = "identity") +
  scale_fill_gradient2(low = my_pal[1], high = my_pal[3], 
                       mid = my_pal[2], midpoint = 0) + 
  theme_minimal() +
  theme(axis.title = element_text(face = "bold", size = 12),
        axis.text.x = 
          element_text(angle = 45, hjust = 1, size =6, face= "bold"),
        axis.text.y = element_text(size =10, face= "bold"),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank()) +
  xlab("MRs") + theme(legend.position = "none") + ylab("Enrichment Score")
```

### S6D. PC explained variance

```{r S6D, warning=FALSE, message=FALSE}
fviz_eig(rna.pr,
         addlabels = F, 
         barcolor = "black", 
         barfill = carto_pal(9, "Vivid")[9], 
         choice = "variance",ncp = 10) + theme_bw()+
  geom_hline(yintercept = 5, linetype="dashed")

```

### S6E. Optimal number of clusters

```{r S6E, warning=FALSE, message=FALSE}
fviz_nbclust(rna.pr$x[,1:3], kmeans, method = "silhouette", linecolor = "black") +
  labs(subtitle = "Silhouette method")

```

### S6F. Clustering and comparison to patient id

```{r S6F, warning=FALSE, message=FALSE}
set.seed(123)
clust <- kmeans(rna.pr$x[,1:3], centers = 2)
clust$cluster <- factor(clust$cluster)
cluster_df <- data.frame(Sample = names(clust$cluster), cluster = clust$cluster)
cluster_df$Patient <- substr(cluster_df$Sample, 1, 3)
cluster_df$cluster <- paste0("C", cluster_df$cluster)
cluster_df_melt <- melt(table(cluster_df$Patient, cluster_df$cluster))
cluster_df_melt$Var.2 <- factor(cluster_df_melt$Var.2)
cluster_df_melt$Var.1 <- factor(cluster_df_melt$Var.1)
colnames(cluster_df_melt) <- c("Patient","Cluster", "value")
ggplot(cluster_df_melt, aes(x=Patient, y=value, fill=Cluster)) + geom_bar(stat="identity", position = "fill") +
  theme_classic2() + scale_fill_manual(values = carto_pal(9, "Vivid")[c(6,7)]) +
  ylab("Fraction") + theme(legend.position = "top") +theme(axis.text.x = element_text(angle=90, vjust = 0.5))
```

## Figure 4

### 4A. PCA colored by clusters (C1/C2)

```{r 4A, warning=FALSE, message=FALSE}
pca_df <- merge(pca_df, cluster_df, by="Sample")
ggplot(pca_df, aes(x = PC1, y = PC2, col=cluster, fill=cluster)) + geom_point(size=4, shape=21, col="black") +
  theme_classic2() + theme(legend.position = "top", legend.title = element_blank()) +
  scale_color_manual(values = carto_pal(9, "Vivid")[c(6,7)],) +
  scale_fill_manual(values = carto_pal(9, "Vivid")[c(6, 7)],) +
  ggforce::geom_mark_ellipse(aes(group=cluster))
```

### 4B. PDO drug screen results

```{r 4B, warning=FALSE, message=FALSE}
viability <- read.table("PC_organoid_drug_screen FC_viability_alldata_hetscore.txt", header = T, sep = "\t", dec = ",") 
rownames(viability) <- paste0(viability$Patient, viability$Core) 
viability <- viability[,-ncol(viability)] 

for(f in 3:ncol(viability)){
  viability[which(viability[,f] > 1),f] <- 1
  #viability[,f] <- scale(viability[,f])
} 

viability$Sample <- substr(rownames(viability), 1, 4) 
viability[which(viability[,f] > 1),f] <- 1
viability <- melt(viability, id.vars = c("Patient", "Core", "Sample"))
viability$variable <- sapply(strsplit(as.character(viability$variable), split = "\\."), "[", 1)  
ggplot(viability, aes(x=reorder(variable, -value), y=value)) + geom_boxplot(fill=carto_pal(9, "Vivid")[c(9)]) + coord_flip() +
  theme_classic2() + geom_jitter(alpha=0.2) +
  ylab("Viablility") + xlab("Deugs")
```


### 4D. PDO drug screen vs clusters (significant results)

```{r 4D, warning=FALSE, message=FALSE}
cluster_df2 <- merge(cluster_df, viability, by="Sample")
cluster_df2 <- merge(cluster_df2, cores, by="Sample")
cluster_df3 <- filter(cluster_df2, variable %in% c("Crizotinib", "Daunorubicin", "Ponatinib", "Sunitinib"))
cluster_df3$variable <- factor(cluster_df3$variable, levels = c("Crizotinib", "Ponatinib", "Daunorubicin", "Sunitinib"))
ggplot(cluster_df3, aes(x=cluster, y=value, fill=cluster)) + geom_boxplot() + 
  stat_compare_means(vjust = 1.5,aes(label = paste0("p = ", after_stat(p.format)))) + theme_classic2() +
  geom_jitter(alpha=0.2) + xlab("Clusters") + ylab("Viability") +
  facet_wrap(~variable, scale="free", nrow = 1) +
  scale_fill_manual(values = carto_pal(9, "Vivid")[c(6, 7)]) +
  theme(legend.position = "none")  + ylim(0,1.01) 
```


## Sup. Figure 8

### S8A. RTK ligands and receptors expression

```{r S8A, warning=FALSE, message=FALSE}
design3 <- merge(design2, cluster_df, by="Sample")
design3$cluster <- factor(design3$cluster)
rownames(design3) <- design3$Sample
df2 <- df[,colnames(df) %in% rownames(design3)]
design3 <- design3[colnames(df2),]

dds <- DESeqDataSetFromMatrix(countData=df2, 
                              colData=design3, 
                              design=~cluster, tidy = FALSE)

dds <- dds[(rowSums(counts(dds) > 3) > 3 ),]
dds <- DESeq(dds)

res <- results(dds, contrast = c("cluster", "C1", "C2"))
res <- as.data.frame(res)
res <- res[order(res$padj, decreasing = F),]
head(res, 50)
res <- cbind.data.frame(res, Gene_name = rownames(res))
res$log2FoldChange[which(res$log2FoldChange < -7)] <- -7
res$log2FoldChange[which(res$log2FoldChange > 7)] <- 7
res$DE <- "ns"
res$DE[which((res$log2FoldChange < -1)&(res$padj<0.01))] <- "Downregulated"
res$DE[which((res$log2FoldChange > 1)&(res$padj<0.01))] <- "Upregulated"
res <- na.omit(res)
my_pal2 <- carto_pal(9, "Vivid")[c(7, 9, 6)]

RTKs <- read.table("Receptors.csv", header = T, sep=",")
res_RTK <- res[RTKs$Approved.symbol,]

RTK_ligands <- read_xlsx("Ligands.xlsx")
res_RTK_l <- res[RTK_ligands$Ligands,]

res_RTK$type="Receptors"
res_RTK_l$type="Ligands"
res_RTK_lr <- rbind(res_RTK, res_RTK_l)
ggplot(res_RTK_lr[which(res_RTK_lr$padj < 0.05),], aes(x = reorder(Gene_name, log2FoldChange), y = log2FoldChange)) + 
  geom_bar(aes(fill = log2FoldChange), stat = "identity") +
  scale_fill_gradient2(low = my_pal2[1], high = my_pal2[3], 
                       mid = my_pal2[2], midpoint = 0) + 
  theme_classic2() +
  theme(axis.title = element_text(face = "bold", size = 12),
        axis.text.x = 
          element_text(angle = 45, hjust = 1, size =8, face= "bold"),
        axis.text.y = element_text(size =10, face= "bold"),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank()) +
  theme(legend.position = "none") + ylab("log2FoldChange (Cluster 1 vs 2)") +
  facet_wrap(~type, nrow = 2, scales = "free") +theme(axis.title.x = element_blank())
```

### S8B. PDO drug screen vs clusters (not significant results)

```{r S8B, warning=FALSE, message=FALSE}
ggplot(cluster_df2[which(cluster_df2$variable %!in% c("Crizotinib", "Daunorubicin", "Ponatinib", "Sunitinib")),], aes(x=cluster, y=value, fill=cluster)) + geom_boxplot() + 
  stat_compare_means(vjust = 1.5,hjust=-1,aes(label = paste0("p = ", after_stat(p.format)))) + theme_classic2() +
  geom_jitter(alpha=0.2) + xlab("Clusters") + ylab("Viability") +
  facet_wrap(~variable, scale="free", ncol = 1) +
  scale_fill_manual(values = carto_pal(9, "Vivid")[c(6, 7)]) +
  theme(legend.position = "none") + ylim(0,1.01)
```

## Figure 5

### 5A. Pathway activity comparison

```{r 5A, warning=FALSE, message=FALSE}
prog <- get_progeny()
colnames(prog)[3] <- "mor"
contrast_acts <- run_viper(t(df_norm), prog, minsize = 5)
contrast_acts <- merge(contrast_acts, cluster_df, by.x="condition", by.y="Sample")
ggplot(contrast_acts[which(contrast_acts$source %in% c("Androgen", "Hypoxia", "NFkB", "p53", "PI3K", "TNFa", "Trail", "WNT")),], 
       aes(x=cluster, y=scale(score, center = 0), fill=cluster)) + geom_boxplot() + 
  stat_compare_means(vjust = 1,aes(label = paste0("p = ", after_stat(p.format)))) + theme_classic2() +
  facet_wrap(~source, scales = "free", nrow = 2) + geom_jitter(alpha=0.2) +
  scale_fill_manual(values = carto_pal(9, "Vivid")[c(6, 7)]) +
  theme(legend.position = "none") +
    ylab("Activity (VIPER)") + xlab("Clusters")
```

### 5B. Machine learning, cross validation plotting and selection of the best model


```{r 5B, warning=FALSE, message=FALSE}
contrast_acts2 <- contrast_acts %>% pivot_wider(id_cols = condition, names_from = source, values_from = score) %>% as.data.frame()
contrast_acts2 <- merge(contrast_acts2, cluster_df, by.x="condition", by.y="Sample")
data <- contrast_acts2
data <- data[,-c(1, ncol(data))]
set.seed(123)
#Put 3/4 of the data into the training set
data_split <- initial_split(data, prop = 3/4, strata = cluster)

#Create df for 2 datasets
train_data <- training(data_split)
test_data <- testing(data_split)


#Feature selection on training dataset
train_data_selected <- train_data
# Define resampling method (cross-validation)
set.seed(345)
train_cv_folds <- vfold_cv(train_data_selected, v = 10, strata = cluster)

#Apply feature selection to the test data
test_data_selected <- test_data


#Create recipe
formula <- as.formula(cluster ~ .)

basic_rec <- recipe(formula, data = train_data)
prep(basic_rec)

#Specify models
log_spec <-
  multinom_reg(penalty = tune(), mixture = 1) %>% 
  set_engine("glmnet") %>%
  set_mode("classification")

rf_spec <- 
  rand_forest( mtry = tune(), min_n = tune(), trees = 100) %>%
  set_engine("ranger", importance = "impurity") %>% # impurity: provide variable importance scores for this model
  set_mode("classification")

xgb_spec <- 
  boost_tree(trees = 1000, tree_depth = tune(), learn_rate = tune()) %>%
  set_engine("xgboost") %>%
  set_mode("classification") 

knn_spec <- 
  nearest_neighbor(neighbors = tune()) %>% 
  set_engine("kknn") %>%
  set_mode("classification") 

svm_spec <-
  svm_rbf(cost = tune(), rbf_sigma = tune()) %>%
  set_mode("classification") %>%
  set_engine("kernlab", maxiter = 10000, tol = 1e-5)

all_workflows <- 
  workflow_set( preproc = list("basic" = basic_rec),
                models = list(logistic = log_spec, random_forest = rf_spec,
                              xgboost = xgb_spec, knn = knn_spec,
                              svm = svm_spec))
#Tuning
#Define grid
log_grid <- grid_regular(
  penalty(), # Penalty (lambda) on log10 scale ,# Mixture (alpha) from 0 (ridge) to 1 (lasso)
  levels = 50  # Number of levels for each parameter
)


num_predictors <- ncol(train_data) - 1 #subtract sample, value and responder

rf_grid <- grid_regular(
  mtry(range = c(1, num_predictors)),   
  min_n(range = c(1, 10)), # min_n between 1 and 10
  levels = 5              # 5 levels for each parameter
)

xgb_grid <- grid_regular(
  tree_depth(range = c(1, 10)),
  learn_rate(range = c(-3, -1)), # log scale for learning rate
  levels = 5
)

knn_grid <- grid_regular(
  neighbors(range = c(1, 20)),
  levels = 10
)

svm_grid <- grid_regular(
  cost(range = c(-5, 2)), # log scale for cost
  rbf_sigma(range = c(-3, 0)), # log scale for sigma
  levels = 5
)


#Tune 
all_workflows <- all_workflows %>%
  option_add(id="basic_logistic", resamples = train_cv_folds,
             grid = log_grid) %>%
  option_add(id="basic_random_forest", resamples = train_cv_folds, 
             grid = rf_grid) %>%
  option_add(id="basic_xgboost", resamples = train_cv_folds, 
             grid = xgb_grid) %>%
  option_add(id="basic_knn", resamples = train_cv_folds, 
             grid = knn_grid) %>%
  option_add(id="basic_svm", resamples = train_cv_folds, 
             grid = svm_grid)

doParallel::registerDoParallel()  
all_results <- all_workflows %>%
  workflow_map(fn = "tune_grid" , resamples = train_cv_folds, 
               verbose = TRUE)


# Check results
roc_results <- rank_results(all_results, rank_metric = "roc_auc")

raw_roc_auc_plot <- autoplot(all_results, metric = "roc_auc") 
#raw_roc_auc_plot + theme_bw() + scale_color_manual(values = pals::kelly()[c(3,4,5,6,8,9)]) +
#  guides(
#    shape = "none") + theme(legend.position = "top") +guides(color=guide_legend(nrow=3,byrow=TRUE))


best_roc_auc_plot <- autoplot(
  all_results,
  rank_metric = "roc_auc",  # <- how to order models
  metric = "roc_auc",       # <- which metric to visualize
  select_best = TRUE     # <- one point per workflow
) +
  lims(y = c(0.5, 1)) +
  theme(
    axis.title = element_text(size = 18),   # Increase axis title size
    axis.text = element_text(size = 14),       # Increase legend text size
    legend.title = element_text(size = 14),
    legend.text = element_text(size = 14))
#best_roc_auc_plot + theme_bw() + scale_color_manual(values = pals::kelly()[c(3,4,5,6,8,9)]) +
#  guides(
#    shape = "none") + theme(legend.position = "top") +guides(color=guide_legend(nrow=3,byrow=TRUE))

raw_accuracy_plot <- autoplot(all_results, metric = "accuracy", select_best = T, std_errs = 1) 
raw_accuracy_plot + theme_bw() + scale_color_manual(values = pals::kelly()[c(3,7,4,5,6,9)]) +
  guides(
    shape = "none") + theme(legend.position = "top") +guides(color=guide_legend(nrow=3,byrow=TRUE)) +
  ylim(0.5,1) 


# Get all workflow IDs
workflow_ids <- all_results %>% pull(wflow_id)

# Loop through all workflows to get the best result based on ROC AUC
best_results_list <- lapply(workflow_ids, function(id) {
  # Extract results for the current workflow ID
  result <- extract_workflow_set_result(all_results, id = id)
  # Get the best result based on ROC AUC
  best_result <- show_best(result, metric = "roc_auc", n = 1)
  # Add workflow_id to the result
  best_result$workflow_id <- id
  return(best_result)
})

# Combine all best results into a single data frame
best_results_df <- bind_rows(best_results_list)

best_model_info <- head(roc_results, 1)

best_model <- all_results %>%
  extract_workflow_set_result(best_model_info$wflow_id) %>%
  select_best(metric = "roc_auc")

# Finalize the model 
final_wflow <- all_results %>%
  extract_workflow(best_model_info$wflow_id) %>% 
  finalize_workflow(best_model)
```


### 5C. Testing best model in unseen data

```{r 5C, warning=FALSE, message=FALSE}
final_fit <- fit(final_wflow, data = train_data_selected)

test_results <- predict(final_fit, new_data = test_data_selected, type = "class") %>%
  bind_cols(test_data_selected)

test_results_prob <- predict(final_fit, new_data = test_data_selected, type = "prob") %>%
  bind_cols(test_data_selected)

# Collect predictions and make confusion matrix
test_results <- test_results %>%
  mutate(
    cluster = as.factor(cluster),
    .pred_class = as.factor(.pred_class)
  )

# Calculate ROC AUC
test_results_prob <- test_results_prob %>%
  mutate(
    cluster = as.factor(cluster))

roc_auc_metric_validation <- test_results_prob %>%
  roc_auc(truth = cluster, .pred_C1)

classification_metrics <- test_results %>%
  metrics(truth = cluster, estimate = .pred_class)

conf_matrix <- test_results %>% 
  conf_mat(truth = cluster, estimate = .pred_class)

#Final visualization and plots ------------------
# Convert confusion matrix to a tidy format
conf_matrix_tidy <- as.data.frame(conf_matrix$table)

# Extract classification metrics (e.g., accuracy, precision, recall, etc.)
accuracy <- classification_metrics %>%
  filter(.metric == "accuracy") %>%
  pull(.estimate)

kap <- classification_metrics %>%
  filter(.metric == "kap") %>%
  pull(.estimate)


roc_auc <- roc_auc_metric_validation %>%
  pull(.estimate)

conf_matrix_heatmap <- ggplot(conf_matrix_tidy, aes(x = Prediction, y = Truth, fill = Freq)) +
  geom_tile() +
  geom_text(aes(label = Freq), color = "white", size = 5) +
  scale_fill_gradientn(colours = rev(pals::ocean.ice(11)[3:11])) +
  labs(
    title = paste0("Confusion Matrix Heatmap"),
    x = "Predicted",
    y = "Actual",
    subtitle = paste0("ROC AUC: ", round(roc_auc, 3), 
                      "\nAccuracy: ", round(accuracy, 3),
                      "\nKap: ", round(kap, 3))
  ) + theme_minimal() +
  theme(axis.title = element_text(size = 16),   # Increase axis title size
        axis.text = element_text(size = 14),
        plot.title = element_text(size = 16, face = "bold"),
        plot.subtitle = element_text(size = 16) )

conf_matrix_heatmap
```


